---
editor_options: 
  markdown: 
    wrap: 72
---

SECTION 4.2.2: Normal model (w/out threshold)
(Section 4.2.3 is also included, it is enough to move the threshold (line 90-91) to 1000)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
rm(list=ls())
```

```{r}
setwd("C:/Users/Simone/Downloads")

library(rjags)
library(caret)
library(ggmcmc)

options(scipen = 999)
```

```{r}
rm(list = ls())

CO2 <- read.csv("C:/Users/Simone/Downloads/CO2.csv")

#Eliminate Luxembourg from data, since it's an outlier for both GDP and CO2 and it represent a unique case that we aren't able to explain with our data
q=quantile(CO2$GDP,0.988)
p=quantile(CO2$co2percap, 0.988)
CO2 <- subset(CO2, CO2$GDP<q)
CO2 <- subset(CO2, CO2$co2percap<p)

```

This model is built with
uninformative flat priors, using the threshold of 27000 to separe in two
groups the high-GDP and low-GDP countries. The observations are modeled
using a simple normal model, where only the mean is influenced by the 4
covariates (even though urbanization and internet level did not belong
to the best selected model, we tried several options and we included
them in the model, to show how their results differ from the covariates
that are mode useful).

```{r}
model_string <- "
model {
  for (i in 1:N) {
    
    # Define the linear predictors, only for mu
    mu[i] <- inprod(X[i,], beta)
    
    #mu[i] <- beta_poor *X[i,1] +
    #         beta_rich *X[i,2] +
    #         beta_poor_gdp *X[i,3] +
    #         beta_rich_gdp *X[i,4] +
    #         beta_poor_EnergyUse *X[i,5] +
    #         beta_rich_EnergyUse *X[i,6] +
    #         beta_poor_urb *X[i,7] +
    #         beta_rich_urb *X[i,8] +
    #         beta_poor_internet *X[i,9] +
    #         beta_rich_internet *X[i,10] +

    #Normal model where the mean is modeled using the linear predictor, without further trasformations, for high interpretability
    co2percap[i] ~ dnorm(mu[i], beta_gamma)

    indicator[i] <- (unGDP[i] >= threshold)
    X[i, 1]   <-                (1 - indicator[i])
    X[i, 2]   <-                indicator[i]
    X[i, 3]   <- GDP[i]       * (1 - indicator[i])
    X[i, 4]   <- GDP[i]       * indicator[i]
    X[i, 5]   <- EnergyUse[i] * (1 - indicator[i])
    X[i, 6]   <- EnergyUse[i] * indicator[i]
    X[i, 7]   <- urb[i]       * (1 - indicator[i])
    X[i, 8]   <- urb[i]       * indicator[i]
    X[i, 9]   <- internet[i]  * (1 - indicator[i])
    X[i, 10]  <- internet[i]  * indicator[i]
    
  }

  #Prior for the variance
  beta_gamma ~ dgamma(0.001, 0.001)
  
  # Priors for beta parameters, beta[4] is the informative
  beta[1] ~ dnorm(0, 0.01)   #beta_poor
  beta[2] ~ dnorm(0, 0.01)   #beta_rich
  beta[3] ~ dnorm(1, 1)      #beta_poor_gdp
  beta[4] ~ dnorm(0, 1)      #beta_rich_gdp
  beta[5] ~ dnorm(0, 0.01)   #beta_poor_EnergyUse
  beta[6] ~ dnorm(0, 0.01)   #beta_rich_EnergyUse
  beta[7] ~ dnorm(0, 0.01)   #beta_poor_urb
  beta[8] ~ dnorm(0, 0.01)   #beta_rich_urb
  beta[9] ~ dnorm(0, 0.01)   #beta_poor_internet
  beta[10]~ dnorm(0, 0.01)   #beta_rich_internet

  # Prior for threshold
  threshold ~ dunif(min_threshold, max_threshold)

  # Define thresholds
  min_threshold <- 27000
  max_threshold <- 27001
}
"

#Normalization of covariates
preproc2 <- preProcess(CO2[,c(4:10)], method=c("range"))
norm2 <- predict(preproc2, CO2[,c(4:10)])

# Data preparation
data_list <- list(
  N = nrow(CO2),    # Number of data points
  co2percap = CO2$co2percap,
  EnergyUse = norm2$EnergyUse * (1-CO2$Lowcarbon_energy/100),
  GDP = norm2$GDP,
  pop = norm2$pop,
  internet = CO2$internet/CO2$pop,
  urb = CO2$urb/100,
  unGDP = CO2$GDP   #Un-normalized GDP to be compared with threshold
)

# Set the number of burn-in and iterations
n_iterations <- 5000

# Compile the model
jags_model <- jags.model(textConnection(model_string), data = data_list)

update(jags_model, 5000, progress.bar="none")

# Run the model
jags_output <- coda.samples(model = jags_model, variable.names = c(
  "beta",
  "threshold",
  "beta_gamma"

), n.iter = n_iterations, n.thin = 4, n.chains = 1)

gr=summary(jags_output)
gr

```
```{r}
mcmc_trace(jags_output,pars = c("beta[1]"))
mcmc_areas(jags_output,pars = c("beta[1]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[2]"))
mcmc_areas(jags_output,pars = c("beta[2]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[3]"))
mcmc_areas(jags_output,pars = c("beta[3]"),  prob = 0.80)
mcmc_trace(jags_output,pars = c("beta[4]"))
mcmc_areas(jags_output,pars = c("beta[4]"),  prob = 0.80)
mcmc_trace(jags_output,pars = c("beta[5]"))
mcmc_areas(jags_output,pars = c("beta[5]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[6]"))
mcmc_areas(jags_output,pars = c("beta[6]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[7]"))
mcmc_areas(jags_output,pars = c("beta[7]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[8]"))
mcmc_areas(jags_output,pars = c("beta[8]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[9]"))
mcmc_areas(jags_output,pars = c("beta[9]"),  prob = 0.95)
mcmc_trace(jags_output,pars = c("beta[10]"))
mcmc_areas(jags_output,pars = c("beta[10]"),  prob = 0.95)

```



$$
\begin{align*}
\mu_i &\leftarrow \text{inprod}(X_i, \beta) \\
\text{co2percap}_i &\sim \mathcal{N}(\mu_i, \beta_\gamma) \\
\text{indicator}_i &\leftarrow (\text{unGDP}_i \geq \text{threshold}) \\
X_{i, 1} &\leftarrow (1 - \text{indicator}_i) \\
X_{i, 2} &\leftarrow \text{indicator}_i \\
X_{i, 3} &\leftarrow \text{GDP}_i \cdot (1 - \text{indicator}_i) \\
X_{i, 4} &\leftarrow \text{GDP}_i \cdot \text{indicator}_i \\
X_{i, 5} &\leftarrow \text{EnergyUse}_i \cdot (1 - \text{indicator}_i) \\
X_{i, 6} &\leftarrow \text{EnergyUse}_i \cdot \text{indicator}_i \\
X_{i, 7} &\leftarrow \text{urb}_i \cdot (1 - \text{indicator}_i) \\
X_{i, 8} &\leftarrow \text{urb}_i \cdot \text{indicator}_i \\
X_{i, 9} &\leftarrow \text{internet}_i \cdot (1 - \text{indicator}_i) \\
X_{i, 10} &\leftarrow \text{internet}_i \cdot \text{indicator}_i \\
\end{align*}
$$ Mean SD Intercept_poor 0.2955 0.32209 Intercept_rich 4.2510 0.70622

The results partially confirm our initial thesis: the parameter
associated to the GDP of low-GDP countries shows a direct
proportionality to the CO2 per capita, while the relation is near zero
for the high-GDP group, even though the zero value is included also in
the 95% credible region of the parameter associated to the GDP of
low-GDP countries (in the picture the 80% credible region is used, to
keep outside the zero axis and show that it is still unprobable that the
covariate is irrelevant). It is also visible how EnergyUse is by far the
covariate that explains the largest part of the variability in the
dataset, indeed it is more than an order of magnitude bigger than the
others both over and under the threshold (this can be a sign of
relevance since all covariate have been normalized). Regarding the
Urbanization and Internet covariate, they are as expected near zero,
with the exception of Urbanization over the GDP threshold: this should
inversely relate the Urbanization of a high-GDP country with its
pollution, this could mean that living in a city could be more efficient
from the pollution point of view. Finally the intercept parameter is
higher for the countries over the GDP threshold, this was expected by
our thesis since it takes on itself the CO2 per capita portion that was
explained by the GDP covariate in the data points under the threshold.


